\chapter{Understanding the Business Domain}


\section{Why understand the business}
Every modern business is a technology business. For some it is the core model. For others, technology supports the core model.
Irrespective of the core business model technology is an essential part of business.
The goal of IT teams is to help the business achieve the business goals, hence IT teams must understand the business model.
Understanding of the business model by IT teams will lead to faster delivery of value to the business.
It will lead to more active participation of the IT teams in the business decisions and it will align the IT objectives with the business objectives.
The most important outcome of having teams of having teams understand the business is that business start to think of it as a trusted partner rather than just a service provider.
What this essentially means is more value to the business, move value to the customers of the business.

\paragraph{important}
Architects and developers etc who understand the business earn the trust of their peers.
I.e don't just learn new techologies, spend time understanding the business.


\section{Introduction to Business Model Canvas}
Business model canvas is a tool that helps discuss, communicate, design and understand the organisations business model.
The benefit of the business model canvas is that the entire business model may be depicted in one image.

To create the business model canvas, one has to focus on nine core areas of the business and these are organised in the form of a canvas.

Do own research

\subsection{Building blocks}
- Nine building blocks
- The first building block is customer segments. Customers are the reason the business exists, so one has to think carefully about who are the customers.
- The next building block is value proposition. What kind of value are we providing to each of these customers.
- The next one is key resources. Emphasis on key, there are many resources that our business requires but you need to think about those resources which are essential for value proposition.
- There may be multiple resources that the business needs, but think about the key ones which without the business cannot exist.
- Next one is the key partners who are suppliers of the key resources to the business.
- Next one is technology providers. Consider procuring technology from other suppliers or partners.
- Also need to consider appropriate permissions to operate. Without them a business will not be able to operate.
- Businesses need to carry out multiple activities. Underkey activities, you need to think about the activities that the business carries out to create the value for the customer.

- Customer retention is one of the most important things for any business, and to retain a customer you need to ensure that the customers are happy with the services you are providing and the relationship you have with them.
- Under the customer relationships, one has to think about the type of relationship that is offered to each of the customer segments.
- The cashflow revenue stream depicts the inflow of revenue for the business for what does.
- The cost structure has the depiction of the cash outflow. These are the expenses incurred by the business on carrying out the key activities.
- Next one is the channels by which the customers would like to be reached.

%TODO insert example canvas diagram here